---
title: "Create dataset"
author: "Nathan Hawkins"
date: "5/17/2021"
output: html_document
---

# This script you should be able to run and get the dataset we need with chulls

# First we read in all the tracking data

```{r}
rm(list = ls())

library(tidyverse)
library(vroom)
file.tracking <- "https://raw.githubusercontent.com/Schmoegurt/Big-Data-Bowl/master/Data/tracking_gameId_2017091713.csv"
tracking.example <- read_csv(file.tracking)

file.game <- "https://raw.githubusercontent.com/nfl-football-ops/Big-Data-Bowl/master/Data/games.csv"
games.sum <- read_csv(file.game) 

file.plays <- "https://raw.githubusercontent.com/nfl-football-ops/Big-Data-Bowl/master/Data/plays.csv"
plays.sum <- read_csv(file.plays)
games_vec <- unique(games.sum$gameId)

#Add in Player data
file.players <- "https://raw.githubusercontent.com/nfl-football-ops/Big-Data-Bowl/master/Data/players.csv"
players.sum <- read_csv(file.players)


# for(i in 1:length(games_vec)) {                             
#     assign(paste0("tracking", games_vec[i]),                                  
#          vroom(paste0("https://raw.githubusercontent.com/ryurko/Big-Data-Bowl/master/Data/tracking_gameId_",games_vec[i],".csv")))
#   print(i)
# }

# my.path is a list of all the games url's.
my.path <- c("https://raw.githubusercontent.com/Schmoegurt/Big-Data-Bowl/master/Data/tracking_gameId_2017090700.csv")
my.data <- list()

# Populate my.path
for(i in 2:length(games_vec)){
  my.path <- c(my.path, paste0("https://raw.githubusercontent.com/ryurko/Big-Data-Bowl/master/Data/tracking_gameId_",games_vec[i],".csv"))
}

# vroom all the my.path entries into a list
for (i in 1:length(my.path)){
    my.data[[i]] <- vroom(my.path[i])
    print(i)
}


library(dplyr)
all_games <- list()

# Trim the list to only include bal_snap, otherwise your computer will break.
for(i in 1:length(games_vec)){
  all_games[[i]] <- my.data[[i]] %>% filter(event == "ball_snap" | displayName == "football")
}

library(data.table)
#rbind the list into a single df
all_games.df <- rbindlist(all_games)

all_games.df <- all_games.df %>% filter(displayName == "football" & frame.id == 1 | event == "ball_snap") %>% arrange(gameId, playId)

rm(all_games)
rm(my.data)
rm(tracking.example)
```

# Merge it with situational data

```{r}
all_games.merged <- all_games.df %>% inner_join(plays.sum) %>% inner_join(games.sum)

# Get rid of special teams plays
all_games.merged <- all_games.merged %>% filter(isSTPlay == FALSE)


# Get rid of duplicate rows
all_games.merged <- all_games.merged[!duplicated(all_games.merged)]

```


# Create an "offense" variable and game_play variable to differentiate between games that have the same play number.

```{r}
# Create an offense/defense variable
# 1 if on offense, 0 for defense
all_games.merged$pos_team <- ifelse(all_games.merged$homeTeamAbbr == all_games.merged$possessionTeam, "home", "away")
all_games.merged$offense_team <- ifelse(all_games.merged$pos_team == all_games.merged$team, 1, 0)
all_games.merged$game_play <- paste0(all_games.merged$gameId, all_games.merged$playId)

```



# Make other variables
# This chunk makes the x, y, and total distance from football variables.
# Makes Box variable.


```{r}
# # This gets me the 22 players when the ball is snapped for non-special teams plays
# normal_plays <- merged.data %>% filter(isSTPlay == FALSE) %>% filter(event == "ball_snap")

# Find the football
football <- all_games.merged %>% filter(isSTPlay == FALSE) %>% filter(displayName == "football")

# Make frame with football x coordinate
football_x <- football[football$displayName== "football" & football$frame.id == 1, c("x", "y", "playId", "gameId", "game_play")]
football_x <- unique(football_x)

# Rename line of scrimmate
names(football_x)[1] <- "football_scrimmage"
names(football_x)[2] <- "football_hash"

length(unique(football_x$game_play))
length(unique(all_games.merged$game_play))
  
# Join 'em together
normal_plays_scrimmage <- all_games.merged %>% inner_join(football_x)

# Find the distance between players and line of scrimmage

normal_plays_scrimmage$dist_x <- normal_plays_scrimmage$football_scrimmage - normal_plays_scrimmage$x
normal_plays_scrimmage$dist_y <- normal_plays_scrimmage$football_hash - normal_plays_scrimmage$y
normal_plays_scrimmage$total_distance_from_football <- sqrt(normal_plays_scrimmage$dist_y^2 + normal_plays_scrimmage$dist_x^2)



# Who is in the box?
# If not in the box then I assume that they are in coverage.
normal_plays_scrimmage$box <- ifelse(abs(normal_plays_scrimmage$dist_x) < 5 &
                                       abs(normal_plays_scrimmage$dist_y) < 8 &
                                             normal_plays_scrimmage$offense_team == 0, 1, 0)

# Make offense players negative and defense positive
# If they are on offense and greater than 0 then multiply it by -1 so that it is negative. 
normal_plays_scrimmage$dist_x <- ifelse(normal_plays_scrimmage$dist_x > 0 & normal_plays_scrimmage$offense_team == 1,
                                        normal_plays_scrimmage$dist_x * -1, normal_plays_scrimmage$dist_x)
# If they are on defense and less than 0 then multiply it by negative 1 so that it is positive.
normal_plays_scrimmage$dist_x <- ifelse(normal_plays_scrimmage$dist_x < 0 & normal_plays_scrimmage$offense_team == 0,
                                        normal_plays_scrimmage$dist_x * -1, normal_plays_scrimmage$dist_x)

#View(normal_plays_scrimmage %>% select(box, offense_team, dist_x, dist_y, total_distance_from_football, displayName))
all_games.merged <- normal_plays_scrimmage

#Merge player data
all_games.merged <- all_games.merged %>% inner_join(players.sum)

```

# This chunk makes cool variables to play with.
Chulls: Square footage covered by the defense at ball snap
far_right_off: horizontal distance from football of furthest right player on the offense
far_left_off: " left "
far_right_def: horizontall distance from football furthest right player on the defense 
far_left_def: " left"
furthest_deep: Vertical distance of furthest defender from football
furthest_deep2: Vertical distance of 2nd furthest (deepest) defender from football
depth_corner1: Vertical distance from football of the leftmost defender
depth_corner1: " rightmost "
chulls_off: The square footage covered by the offense at snap of the ball
wr_space: The average distance between the wide receivers and nearest defenders at snape of the ball
ol_rating: Average madden rating of the offensive line

```{r}
#I used what was uploaded in the github
madden17QB <- read.csv('madden17QB.csv')
madden17RB <- read.csv('madden17RB.csv')
#All Madden ratings. Available on github
Madden <- read_csv("../Madden NFL 17 Launch Ratings - Full League.csv")

#Make it merge-able
madden_new_col <- Madden %>% 
  rename(LastName = Last, FirstName = First, PositionAbbr = Position) %>%
  select(LastName, FirstName, PositionAbbr, OVR)

madden_ol <- madden_new_col %>%
  filter(PositionAbbr %in% c("LT", "RT", "C", "RG", "LG"))

madden_ol$PositionAbbr <- "OL"

madden_ol <- madden_ol %>% select(-PositionAbbr) %>% rename(ol_rating = OVR)

#error
ol_chulls <- merge(x = all_games.merged, y = madden_ol, by = c("LastName", "FirstName"), all = TRUE)
qb_chulls <- merge(x = ol_chulls, y = madden17QB, by = "displayName", all = TRUE)
all_games.merged <- merge(x = qb_chulls, y = madden17RB, by = "displayName", all = TRUE)

chulls <- NA
far_right_off <- NA
far_left_off <- NA
far_left_def <- NA
far_right_def <- NA
depth_safety1 <- NA
depth_safety2 <- NA
depth_corner1 <- NA
depth_corner2 <- NA
furthest_deep <- NA
furthest_deep2 <- NA
loop_data <- NA
qbrating <- NA
rbrating <- NA
wr_space <- NA
off_chulls <- NA
chulls_ratio <- NA
olrating <- NA
game <- NA


all_games.merged <- all_games.merged %>% filter(displayName != "football")

for(i in 1:length(unique(all_games.merged$game_play))){
  library(dplyr)
    
   # Get the play I want
    area <- all_games.merged %>% filter(game_play == unique(all_games.merged$game_play)[i]) %>% filter(displayName != "football")
    
    #Should be i instead of 1? 
    
  if(nrow(area) > 21 & !is.na(area$dist_x[1])){
    #Get offense & defense for each play
    defense <- area %>% filter(offense_team == 0) %>% dplyr::select(x,y)
    offense_xy <- area %>% filter(offense_team == 1) %>% dplyr::select(x,y)
    offense <- area %>% filter(offense_team == 1) %>% dplyr::select(x,y, PositionAbbr)
    
    # Find the convex hull
    library(grDevices)
    convex_hull <- defense %>% dplyr::slice(chull(x, y)) 
    off_convex_hull <- offense_xy %>% dplyr::slice(chull(x, y))
  
    
    # Find chulls
    chulls[i] <- geometry::polyarea(convex_hull$x, convex_hull$y)
    off_chulls[i] <- geometry::polyarea(off_convex_hull$x, off_convex_hull$y)
    
    #chulls ratio = offensive area vs defensive area 
    chulls_ratio[i] <- off_chulls[i] / chulls[i]
    
    #Get wide receivers
    wrs <- offense %>% filter(PositionAbbr == "WR")
    
    if (nrow(wrs) > 0) {
      wr_dist <- NULL
      
      for (j in 1:nrow(wrs)) {
        wr_xy <- wrs[j,1:2]
        #combine with defense
        def_wr <- rbind(wr_xy, defense)
        dist_matix <- as.matrix(dist(def_wr, diag = TRUE, upper = TRUE))
        diag(dist_matix) <- Inf
        wr_dist[j] <- apply(dist_matix, 1, min)[1]
      }
      
      wr_space[i] <- mean(wr_dist)
    }
    else {
      wr_space[i] <- NA
    }
    
    # Assign distances
    game[i] <- area$game_play[i]
    far_right_off[i] <- max(area[area$offense_team == 1, "dist_y"])
    far_left_off[i] <- min(area[area$offense_team == 1, "dist_y"])
    far_right_def[i] <- max(area[area$offense_team == 0, "dist_y"])
    far_left_def[i] <- min(area[area$offense_team == 0, "dist_y"])
    furthest_deep[i] <- max(area[area$offense_team == 0, "dist_x"])
    value <- area %>% filter(offense_team == 0) %>% arrange(abs(dist_x)) %>% dplyr::select(dist_x) %>% slice(10)
    furthest_deep2[i] <- value$dist_x
    corner1 <- area[area$offense_team == 0 & area$dist_y == max(far_left_def[i]), "dist_x"]
    depth_corner1[i] <- corner1$dist_x
    corner2 <- area[area$offense_team == 0 & area$dist_y == max(far_right_def[i]), "dist_x"]
    depth_corner2[i] <- corner2$dist_x
    qbrating[i] <- mean(area$QBrating, na.rm = TRUE)
    rbrating[i] <- mean(area$RBrating, na.rm = TRUE)
    olrating[i] <- mean(area$ol_rating, na.rm = TRUE)
  }
    else {
      game[i] <- area$game_play[i]
      chulls[i] <- NA
      far_right_off[i] <- NA
      far_left_off[i] <- NA
      far_left_def[i] <- NA
      far_right_def[i] <- NA
      depth_safety1[i] <- NA
      depth_safety2[i] <- NA
      depth_corner1[i] <- NA
      depth_corner2[i] <- NA
      furthest_deep[i] <- NA
      furthest_deep2[i] <- NA
      loop_data[i] <- NA
      qbrating[i] <- NA
      rbrating[i] <- NA
      wr_space[i] <- NA
      off_chulls[i] <- NA
      chulls_ratio[i] <- NA
      olrating[i] <- NA
    }
}

```





# Merge chulls with the dataset that has everything. 

```{r}
chulls_df <- data.frame(chulls = chulls, game_play = unique(all_games.merged$game_play), off_right = far_right_off, off_left = far_left_off, def_right = far_right_def, def_left = far_left_def, def_deep_1 = furthest_deep, def_deep2 = furthest_deep2, corner_right_deep = depth_corner1, corner_left_deep = depth_corner2, QBrating = qbrating, RBrating = rbrating, OLrating = olrating, WR_space = wr_space, chulls_off = off_chulls, chulls_ratio = chulls_ratio)


chulls_merge <- all_games.merged %>% inner_join(chulls_df, by = "game_play")

chulls_merge <- chulls_merge[!is.na(chulls_merge$corner_left_deep),]

```


# Merge in PBP

```{r}
library(nflfastR)
pbp2017 <- nflfastR::load_pbp(2017)

names(pbp2017)[1] <- "playId"
names(pbp2017)[3] <- "gameId"

pbpPlays2017 <- load_pbp(2017)


# Change column names in pbp to make it easily joinable to allPlays2017
colnames(pbpPlays2017)[3] <- "gameId"
colnames(pbpPlays2017)[1] <- "playId"
#Change gameId in pbp data to a double
pbpPlays2017$gameId <- as.double(pbpPlays2017$gameId)

# Join the datasets together
chulls_pbp <- left_join(chulls_merge, pbpPlays2017, 
                        by = c("gameId", "playId"))

```


# How to do some graphics

```{r}

# Make sure you use chulls.model. This gives you just one instance of each play. Otherwise your results will be messed up.
chulls.model <- chulls_pbp %>% arrange(game_play) %>% group_by(game_play) %>% slice(1)

library(ggplot2)
ggplot(data = chulls.model, mapping = aes(x = chulls)) + 
  geom_histogram()


ggplot(data = chulls.model, mapping = aes(x = chulls, y = as.factor(pass))) + 
  geom_boxplot() + 
  facet_wrap(~down.y)
```
```{r}
library(dplyr)

pass_chull <- glm(data = chulls.model %>% filter(down.y < 4) %>% filter(quarter < 4) %>% filter(yardline_100 > 40), formula = as.factor(pass) ~ as.factor(down.y) + yardsToGo +  yardline_100 + chulls + defendersInTheBox, family = binomial(link = "logit"))



# Cool model results. Not robust yet, this needs a lot of work.


result_chull <- glm(data = chulls.model %>% filter(down.y < 4) %>% filter(quarter < 4) %>% filter(yardline_100 > 20), formula = PlayResult ~ as.factor(down.y) + yardsToGo +  yardline_100 + chulls)

summary(result_chull)
```

```{r}
plot(chulls.model$chulls, chulls.model$off_right - chulls.model$off_left)
```


```{r}
chulls.model$safety_depth <- (chulls.model$def_deep2 + chulls.model$def_deep_1)/2
chulls.model$avg_corner_depth <- (chulls.model$corner_left_deep + chulls.model$corner_right_deep)/2
chulls.model$off_spread <- abs(chulls.model$off_left) + abs(chulls.model$off_right)

pass_chull <- glm(data = chulls.model %>% filter(down.y < 4) %>% filter(quarter < 4) %>% filter(yardline_100 > 40), formula = as.factor(pass) ~ as.factor(down.y) + yardsToGo +  yardline_100 + chulls + defendersInTheBox + safety_depth + avg_corner_depth, family = binomial(link = "logit"))

summary(pass_chull)
```


```{r}
result_chull <- glm(data = chulls.model %>% filter(down.y < 4) %>% filter(quarter < 3) %>% filter(yardline_100 > 20), formula = PlayResult ~ as.factor(down.y) + yardsToGo +  yardline_100 + chulls + safety_depth + avg_corner_depth * pass)

summary(result_chull)

result_chull <- glm(data = chulls.model %>% filter(quarter < 3) %>% filter(yardline_100 > 20) %>% filter(pass == 1), formula = PlayResult ~ as.factor(down.y) + yardsToGo +  yardline_100 + chulls + safety_depth + avg_corner_depth + off_spread + defendersInTheBox)

summary(result_chull)
```




# Add other variables that may be useful

```{r}
# Get all pass and rush plays from first 6 weeks of 2017 season
databowlPlays2017 <- vroom("https://raw.githubusercontent.com/nfl-football-ops/Big-Data-Bowl/master/Data/plays.csv")
# Get all play by play data from 2017 season
pbpPlays2017 <- load_pbp(2017)
# Change column names in pbp to make it easily joinable to allPlays2017
colnames(pbpPlays2017)[3] <- "gameId"
colnames(pbpPlays2017)[1] <- "playId"
#Change gameId in pbp data to a double
pbpPlays2017$gameId <- as.double(pbpPlays2017$gameId)
# Join the datasets together
plays2017 <- inner_join(databowlPlays2017, pbpPlays2017, 
                        by = c("gameId", "playId"))
```




# Weekly QBR

```{r}
# Adding Weeks 1-6 QBR to dataset
# First, get the data from Nate's file
qbr2017 <- read_csv("../Simple Dataset/2017-2018 Weekly ESPN QBR.csv")
# Get only 2017 season
qbr2017 <- qbr2017 %>% 
  filter(season == 2017)
# Create a qbr column initiated with all zeros for pass plays and NA for run plays
plays2017$qbr <- rep(0, dim(plays2017)[1])
# Gather the QB names from plays 2017
teams <- plays2017 %>% 
  group_by(posteam) %>% 
  filter(!is.na(posteam)) %>% 
  summarize(Count = n())
# Add Weeks 1-6 QBR for each QB to every play in plays2017
for (i in teams[[1]]) {
  for (j in 1:6) {
    for (k in 1:dim(plays2017[plays2017$week == j & plays2017$posteam == i, 12])[1]) {
      plays2017[plays2017$week == j & plays2017$posteam == i, "qbr"][[1]][k] <- qbr2017[qbr2017$team_abb == i & qbr2017$week_text == str_c("Week ", j), 11][[1]][1]
    }
  }
}
```




# Defensive ranking at end of 2017 - not super useful, but in case we wanna
# do something with it

```{r}
# This code will add the defensive ranking to the plays2017 dataset
# Get defensive ranking from 2017
library(rvest)
defRanks <- "https://www.pro-football-reference.com/years/2017/opp.htm" %>% 
  read_html() %>% 
  html_node("table") %>% 
  html_table()
colnames(defRanks) <- c("Rank", "Team", "Games", "PF", "TotalYds", "TotalPlays", "YardsPerPlay", 
               "Takeaways", "FumblesLost", "FirstDowns", "PassComp", "PassAtt", "PassYds",
                 "PassTD", "Int", "YardsPerPass", "PassFirstDowns", "RunAtt", "RunYds", 
                 "RunTD", "YardsPerRun", "RunFirstDowns", "Penalties", "PenaltyYds", 
                 "PenaltyFirstDowns", "OffenseScorePercent", "TakeawayPercent",
                 "ExpectedPoints")
defRanks <- defRanks %>% 
  slice(2:33) 
# Get team abbreviations
teamAbbreviations <- plays2017 %>% 
  group_by(defteam) %>% 
  summarize(Count = n())
# Set up the data to match with the defensive ranking data
# Give defRanks the appropriate team abbreviations
# First, change Oakland to Las Vegas
teamAbbreviations$defteam <- ifelse(teamAbbreviations$defteam == "LV", "OAK", teamAbbreviations$defteam)
# Make teamAbbreviations an atomic vector
teamAbbreviations <- teamAbbreviations[[1]]
# Organize the names in alphabetical order
teamAbbreviations <- teamAbbreviations[order(teamAbbreviations)]
# Switch LA Rams and LA Chargers (helps with alphabetical order later on)
x <- teamAbbreviations[18] 
teamAbbreviations[18] <- teamAbbreviations[17]
teamAbbreviations[17] <- x
teamAbbreviations
# Give defRanks the appropriate team abbreviations
defRanks$TeamAbb <- rep(0, dim(defRanks)[1])
defRanks[order(defRanks[, "Team"]), ]$TeamAbb <- teamAbbreviations
# Now, apply defensive rank to the plays2017 dataset
plays2017$defRank <- rep(0, dim(plays2017)[1])
for (i in 1:dim(plays2017)[1]) {
  for (j in defRanks$TeamAbb) {
    if (plays2017$defteam[i] == j) {
      plays2017$defRank[i] <- defRanks[defRanks$TeamAbb == j, 1]
    }
  }
}
plays2017$defRank <- as.integer(plays2017$defRank)
```




# Adding number of sacks, interceptions and average number of pass rushers
# at every point in the game

```{r}
# First, gather all gameIds
gameIds <- plays2017 %>% 
  group_by(gameId) %>% 
  summarize(Count = n())
gameIds <- gameIds[[1]]
# Add whether or not the defense blitzed on a pass play
plays2017$isBlitz <- ifelse(plays2017$numberOfPassRushers > 4, 1, 0)
for (i in 1:dim(plays2017)[1]) {
  if (is.na(plays2017$isBlitz[i])) {
    plays2017$isBlitz[i] <- 0
  }
}
# Loops through the dataset and inserts the total number of sacks, interceptions,
# and average pass rushers thus far in a game.
plays2017$totalSacksGame <- rep(0, dim(plays2017)[1])
plays2017$totalIntGame <- rep(0, dim(plays2017)[1])
plays2017$avgPassRushers <- rep(0, dim(plays2017)[1])
plays2017$totalPassBlitzGame <- rep(0, dim(plays2017)[1])
for (i in gameIds) {
  for (j in teams[[1]]) {
    totalSacks <- 0
    totalInt <- 0
    totalBlitz <- 0
    avgPassRushers <- 0
    sumPassRushers <- 0
    passRushPlays <- 0
    for (k in 1:dim(plays2017)[1]) {
      if (plays2017$gameId[k] == i & plays2017$defteam[k] == j) {
        # Sacks
        totalSacks <- totalSacks + plays2017$sack[k]
        plays2017$totalSacksGame[k] <- totalSacks
        # Interceptions
        totalInt <- totalInt + plays2017$interception[k]
        plays2017$totalIntGame[k] <- totalInt
        # Blitz
        totalBlitz <- totalBlitz + plays2017$isBlitz[k]
        plays2017$totalPassBlitzGame[k] <- totalBlitz
        if (!is.na(plays2017$numberOfPassRushers[k])) {
          # Average pass rushers
          passRushPlays <- passRushPlays + 1
          sumPassRushers <- sumPassRushers + plays2017$numberOfPassRushers[k]
          avgPassRushers <- sumPassRushers / passRushPlays
          plays2017$avgPassRushers[k] <- avgPassRushers
        } else {
          plays2017$avgPassRushers[k] <- avgPassRushers
        }
      }
    }
  }
}
```



# Defensive personnel on each play

```{r}
# Add number of defensive personnel on any given play
plays2017$numDL <- rep(0, dim(plays2017)[1])
plays2017$numLB <- rep(0, dim(plays2017)[1])
plays2017$numDB <- rep(0, dim(plays2017)[1])
for (i in 1:dim(plays2017)[1]) {
  plays2017$numDL[i] <- substr(plays2017$personnel.defense[i], 1, 1)
  plays2017$numLB[i] <- substr(plays2017$personnel.defense[i], 7, 7)
  plays2017$numDB[i] <- substr(plays2017$personnel.defense[i], 13, 13)
}
plays2017$numDL <- as.integer(plays2017$numDL)
plays2017$numLB <- as.integer(plays2017$numLB)
plays2017$numDB <- as.integer(plays2017$numDB)
```


# Merge these variables with chulls.model

```{r}
chulls.model <- inner_join(chulls.model, plays2017)
colnames(chulls.model)
```


# Add Nate's epa data

```{r}
epa <- read.csv("rolling epa per play.csv")
chulls.model <- inner_join(chulls.model, epa)
write.csv(chulls.model, "FinalDataset.csv")
```
