---
title: "Nate Notebook"
output: html_notebook
---

```{r}
#to get data
library(nflfastR)
library(espnscrapeR)

#other
library(tidyverse)
library(ggplot2)
library(vroom)
library(randomForest)
library(useful)
library(class)
```


## Get nflfastR data
```{r}
pbp_2017 <- nflfastR::load_pbp(seasons = 2017)

pbp_2017 %>% head(15)
```

## Get Big Data Bowl Data
```{r}
#play by play data (looks like it is just a few weeks of data in the 2017 season)
plays <- vroom("https://raw.githubusercontent.com/nfl-football-ops/Big-Data-Bowl/master/Data/plays.csv")

#tracking data for gameid 2017090700 -- may not use this at all
#tracking <- vroom("https://raw.githubusercontent.com/nfl-football-ops/Big-Data-Bowl/master/Data/tracking_gameId_2017090700.csv")

plays %>% head(5)
```


#### Clean plays dataframe
```{r}
#SPLIT PERSONNEL AND ADD TO DATAFRAME

#split string for the defensive personnel
split_def <- stringr::str_split_fixed(plays$personnel.defense, " ", n = 7) %>% 
                          data.frame() %>% 
                          select(X1, X3, X5)
names(split_def) <- c("DL", "LB", "DB")

#split string for the offensive personnel
split_off <- stringr::str_split_fixed(plays$personnel.offense, " ", n = 7) %>% 
                          data.frame() %>% 
                          select(X1, X3, X5)
names(split_off) <- c("RB", "TE", "WR")

#add to plays data and create plays_pers (plays personnel)
plays_pers <- plays %>% cbind(split_def) %>% cbind(split_off)

#view first 20 rows
plays_pers %>% head(20)
```

#### Bind Big Data Bowl and nflfastR data
```{r}
#clean up column names for join compatibility
names <- pbp_2017 %>% names()
names[1] <- "playId"
names[3] <- "gameId"
names(pbp_2017) <- names
pbp_2017$gameId <- as.double(pbp_2017$gameId)

#join data frames
plays_2017 <- inner_join(plays_pers, pbp_2017, 
                        by = c("gameId", "playId"))
```


### Filter out unwanted plays
```{r}
#filter out special teams plays
plays_2017 <- plays_2017 %>% filter(is.na(SpecialTeamsPlayType) == TRUE)

#filter out penalties
plays_2017 <- plays_2017 %>% filter(isPenalty == FALSE)
```

## Possible Classifications:
- Pass
  - short/deep (pass_length)
  - air_yards (possibly bin this off to get more possibilities than above)
  - pass_location
- Run
  - middle/right/left (run_location)
  - run_gap (combine with above variable -- maybe just use this one with middle because the side may not really matter but the gap does)


### Simplify classifications to one variable
```{r}
#get pass description classification
plays_2017$pass_description <- ifelse(is.na(plays_2017$pass_length), 
                                      "", 
                                      paste("pass", 
                                            plays_2017$pass_length, 
                                            plays_2017$pass_location))
#get run description classification
plays_2017$run_description <- ifelse(is.na(plays_2017$run_location), 
                                     "", 
                                     ifelse(plays_2017$run_location == "middle", 
                                            "run middle", 
                                            paste("run", plays_2017$run_gap)))

#combine descriptions
plays_2017$classification <- paste(plays_2017$pass_description, plays_2017$run_description, sep = "")

#view descriptions and end classification
plays_2017 %>% select(pass_description, run_description, classification)
```
### Features to use in model:
- defensive personnel (defendersInTheBox, DL, LB, DB)
- game situation (quarter, down.x, yardToGo, yardline_100)
- numberOfPassRushers


## Get model data
```{r}
#get model data
model_data <- plays_2017 %>% select(classification, defendersInTheBox, DL, LB, DB, quarter, down.x, yardsToGo, yardline_100) %>%
                             filter(classification != "")

### clean features for model data

#change DL, LB, DB into numeric data types
model_data$DL <- as.numeric(model_data$DL)
model_data$LB <- as.numeric(model_data$LB)
model_data$DB <- as.numeric(model_data$DB)

#impute missing values with column median -- consider throwing these rows out
model_data <- useful::simple.impute(model_data, fun = stats::median)

write.csv(model_data, "Model Data.csv")

#should return 0 if there is not missing values/NA's
#which(is.na(model_data))

### Split into training and testing sets
random <- sample(1:nrow(model_data), 
                 .3 * nrow(model_data), 
                 replace = FALSE)

train <- model_data[-random, ]
test <- model_data[random, ]
train_y <- model_data[-random, 1]
test_y <- model_data[random, 1]
```

## Run Models
```{r}
#Random Forest
rf <- randomForest(classification ~ defendersInTheBox + DL + LB + DB + quarter + down.x + yardsToGo + yardline_100,
                   proximity = TRUE,
                        data = train)


#KNN
knn <- class:knn(train, test, cl = train_y, prob = TRUE, k = 10)



```



##### POSSIBLE ADD INS
- offensive formation
- yardlineSide
- rolling epa/play on passes and rushes for offense and defense
- qbr/defensive rating
- more in depth game information (time of the clock, etc.)
- weather




















