---
title: "Create dataset"
author: "Nathan Hawkins"
date: "5/17/2021"
output: html_document
---

# This script you should be able to run and get the dataset we need with chulls

# First we read in all the tracking data

```{r}
rm(list = ls())

library(tidyverse)
library(vroom)
file.tracking <- "https://raw.githubusercontent.com/Schmoegurt/Big-Data-Bowl/master/Data/tracking_gameId_2017091713.csv"
tracking.example <- read_csv(file.tracking)

file.game <- "https://raw.githubusercontent.com/nfl-football-ops/Big-Data-Bowl/master/Data/games.csv"
games.sum <- read_csv(file.game) 

file.plays <- "https://raw.githubusercontent.com/nfl-football-ops/Big-Data-Bowl/master/Data/plays.csv"
plays.sum <- read_csv(file.plays)
games_vec <- unique(games.sum$gameId)



# for(i in 1:length(games_vec)) {                             
#     assign(paste0("tracking", games_vec[i]),                                  
#          vroom(paste0("https://raw.githubusercontent.com/ryurko/Big-Data-Bowl/master/Data/tracking_gameId_",games_vec[i],".csv")))
#   print(i)
# }

# my.path is a list of all the games url's.
my.path <- c("https://raw.githubusercontent.com/Schmoegurt/Big-Data-Bowl/master/Data/tracking_gameId_2017090700.csv")
my.data <- list()

# Populate my.path
for(i in 2:length(games_vec)){
  my.path <- c(my.path, paste0("https://raw.githubusercontent.com/ryurko/Big-Data-Bowl/master/Data/tracking_gameId_",games_vec[i],".csv"))
}

# vroom all the my.path entries into a list
for (i in 1:length(my.path)){
    my.data[[i]] <- vroom(my.path[i])
    print(i)
}


library(dplyr)
all_games <- list()

# Trim the list to only include bal_snap, otherwise your computer will break.
for(i in 1:length(games_vec)){
  all_games[[i]] <- my.data[[i]] %>% filter(event == "ball_snap" | displayName == "football")
}

library(data.table)
#rbind the list into a single df
all_games.df <- rbindlist(all_games)

all_games.df <- all_games.df %>% filter(displayName == "football" & frame.id == 1 | event == "ball_snap") %>% arrange(gameId, playId)

rm(all_games)
rm(my.data)
rm(tracking.example)
```

# Merge it with situational data

```{r}
all_games.merged <- all_games.df %>% inner_join(plays.sum) %>% inner_join(games.sum)

# Get rid of special teams plays
all_games.merged <- all_games.merged %>% filter(isSTPlay == FALSE)


# Get rid of duplicate rows
all_games.merged <- all_games.merged[!duplicated(all_games.merged)]

```


# Create an "offense" variable and game_play variable to differentiate between games that have the same play number.

```{r}
# Create an offense/defense variable
# 1 if on offense, 0 for defense
all_games.merged$pos_team <- ifelse(all_games.merged$homeTeamAbbr == all_games.merged$possessionTeam, "home", "away")
all_games.merged$offense_team <- ifelse(all_games.merged$pos_team == all_games.merged$team, 1, 0)
all_games.merged$game_play <- paste0(all_games.merged$gameId, all_games.merged$playId)

```



# Make other variables
# This chunk makes the x, y, and total distance from football variables.
# Makes Box variable.


```{r}
# # This gets me the 22 players when the ball is snapped for non-special teams plays
# normal_plays <- merged.data %>% filter(isSTPlay == FALSE) %>% filter(event == "ball_snap")

# Find the football
football <- all_games.merged %>% filter(isSTPlay == FALSE) %>% filter(displayName == "football")

# Make frame with football x coordinate
football_x <- football[football$displayName== "football" & football$frame.id == 1, c("x", "y", "playId", "gameId", "game_play")]
football_x <- unique(football_x)

# Rename line of scrimmate
names(football_x)[1] <- "football_scrimmage"
names(football_x)[2] <- "football_hash"

length(unique(football_x$game_play))
length(unique(all_games.merged$game_play))
  
# Join 'em together
normal_plays_scrimmage <- all_games.merged %>% inner_join(football_x)

# Find the distance between players and line of scrimmage

normal_plays_scrimmage$dist_x <- normal_plays_scrimmage$football_scrimmage - normal_plays_scrimmage$x
normal_plays_scrimmage$dist_y <- normal_plays_scrimmage$football_hash - normal_plays_scrimmage$y
normal_plays_scrimmage$total_distance_from_football <- sqrt(normal_plays_scrimmage$dist_y^2 + normal_plays_scrimmage$dist_x^2)



# Who is in the box?
# If not in the box then I assume that they are in coverage.
normal_plays_scrimmage$box <- ifelse(abs(normal_plays_scrimmage$dist_x) < 5 &
                                       abs(normal_plays_scrimmage$dist_y) < 8 &
                                             normal_plays_scrimmage$offense_team == 0, 1, 0)

# Make offense players negative and defense positive
# If they are on offense and greater than 0 then multiply it by -1 so that it is negative. 
normal_plays_scrimmage$dist_x <- ifelse(normal_plays_scrimmage$dist_x > 0 & normal_plays_scrimmage$offense_team == 1,
                                        normal_plays_scrimmage$dist_x * -1, normal_plays_scrimmage$dist_x)
# If they are on defense and less than 0 then multiply it by negative 1 so that it is positive.
normal_plays_scrimmage$dist_x <- ifelse(normal_plays_scrimmage$dist_x < 0 & normal_plays_scrimmage$offense_team == 0,
                                        normal_plays_scrimmage$dist_x * -1, normal_plays_scrimmage$dist_x)

#View(normal_plays_scrimmage %>% select(box, offense_team, dist_x, dist_y, total_distance_from_football, displayName))
all_games.merged <- normal_plays_scrimmage


```

# This chunk makes cool variables to play with.
Chulls: Square footage covered by the defense at ball snap
far_right_off: horizontal distance from football of furthest right player on the offense
far_left_off: " left "
far_right_def: horizontall distance from football furthest right player on the defense 
far_left_def: " left"
furthest_deep: Vertical distance of furthest defender from football
furthest_deep2: Vertical distance of 2nd furthest (deepest) defender from football
depth_corner1: Vertical distance from football of the leftmost defender
depth_corner1: " rightmost "

```{r}
chulls <- NA
far_right_off <- NA
far_left_off <- NA
far_left_def <- NA
far_right_def <- NA
depth_safety1 <- NA
depth_safety2 <- NA
depth_corner1 <- NA
depth_corner2 <- NA
furthest_deep <- NA
furthest_deep2 <- NA
loop_data <- NA

all_games.merged <- all_games.merged %>% filter(displayName != "football")

for(i in 1:length(unique(all_games.merged$game_play))){
  library(dplyr)
    
   # Get the play I want
    area <- all_games.merged %>% filter(game_play == unique(all_games.merged$game_play)[i]) %>% filter(displayName != "football")
    
  if(nrow(area) > 21 & !is.na(area$dist_x[1])){
    defense <- area %>% filter(offense_team == 0) %>% dplyr::select(x,y)
    
    # Find the convex hull
    library(grDevices)
    convex_hull <- defense %>% dplyr::slice(chull(x, y)) 
  
    
    # Find chulls
    chulls[i] <- geometry::polyarea(convex_hull$x, convex_hull$y)
    
    
    # Assign distances
    far_right_off[i] <- max(area[area$offense_team == 1, "dist_y"])
    far_left_off[i] <- min(area[area$offense_team == 1, "dist_y"])
    far_right_def[i] <- max(area[area$offense_team == 0, "dist_y"])
    far_left_def[i] <- min(area[area$offense_team == 0, "dist_y"])
    furthest_deep[i] <- max(area[area$offense_team == 0, "dist_x"])
    value <- area %>% filter(offense_team == 0) %>% arrange(abs(dist_x)) %>% dplyr::select(dist_x) %>% slice(10)
    furthest_deep2[i] <- value$dist_x
    corner1 <- area[area$offense_team == 0 & area$dist_y == max(far_left_def[i]), "dist_x"]
    depth_corner1[i] <- corner1$dist_x
    corner2 <- area[area$offense_team == 0 & area$dist_y == max(far_right_def[i]), "dist_x"]
    depth_corner2[i] <- corner2$dist_x
  }
}

```





# Merge chulls with the dataset that has everything. 

```{r}

chulls_df <- data.frame(chulls = chulls, game_play =  unique(all_games.merged$game_play), off_right = far_right_off, off_left = far_left_off, def_right = far_right_def, def_left = far_left_def, def_deep_1 = furthest_deep, def_deep2 = furthest_deep2, corner_right_deep = depth_corner1, corner_left_deep = depth_corner2)

chulls_merge <- all_games.merged %>% inner_join(chulls_df)

chulls_merge <- chulls_merge[!is.na(chulls_merge$corner_left_deep),]

```


# Merge in PBP

```{r}
library(nflfastR)
pbp2017 <- nflfastR::load_pbp(2017)

names(pbp2017)[1] <- "playId"
names(pbp2017)[3] <- "gameId"

pbpPlays2017 <- load_pbp(2017)


# Change column names in pbp to make it easily joinable to allPlays2017
colnames(pbpPlays2017)[3] <- "gameId"
colnames(pbpPlays2017)[1] <- "playId"
#Change gameId in pbp data to a double
pbpPlays2017$gameId <- as.double(pbpPlays2017$gameId)

# Join the datasets together
chulls_pbp <- left_join(chulls_merge, pbpPlays2017, 
                        by = c("gameId", "playId"))

colnames(chulls_pbp)
```


# How to do some graphics

```{r}

# Make sure you use chulls.model. This gives you just one instance of each play. Otherwise your results will be messed up.
chulls.model <- chulls_pbp %>% arrange(game_play) %>% group_by(game_play) %>% slice(1)

library(ggplot2)
ggplot(data = chulls.model, mapping = aes(x = chulls)) + 
  geom_histogram()


ggplot(data = chulls.model, mapping = aes(x = chulls, y = as.factor(pass))) + 
  geom_boxplot() + 
  facet_wrap(~down.y)
```



# Cool model results. Not robust yet, this needs a lot of work.

```{r}
library(dplyr)

chulls.model$corner_left_deep

pass_chull <- glm(data = chulls.model %>% filter(down.y < 4) %>% filter(quarter < 4) %>% filter(yardline_100 > 40), formula = as.factor(pass) ~ as.factor(down.y) + yardsToGo +  yardline_100 + chulls + defendersInTheBox, family = binomial(link = "logit"))

result_chull <- glm(data = chulls.model %>% filter(down.y < 4) %>% filter(quarter < 4) %>% filter(yardline_100 > 20), formula = PlayResult ~ as.factor(down.y) + yardsToGo +  yardline_100 + chulls)

summary(result_chull)
```
# Add Madden QB and RB ratings

madden17QB <- read.csv('https://raw.githubusercontent.com/cadentwebb/football/main/Simple%20Dataset/madden17QB.csv?token=ASLGDPENNK5BHX555LCR37DAVQZPK')

madden17RB <- read.csv('https://raw.githubusercontent.com/cadentwebb/football/main/Simple%20Dataset/madden17RB.csv?token=ASLGDPBYGWQY3SYYCOTNMKLAVQZSS')


qb_chulls <- merge(x = chulls.model, y = madden17QB, by = "displayName", all = TRUE)
chulls.model <- merge(x = qb_chulls, y = madden17RB, by = "displayName", all = TRUE)



